---
title: "Reproducible Research: Peer Assessment 1"
author: "dillonchewwx"
date: "05/02/2021"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
We will use the tidyverse package to read and process the data.
We will use ggpubr for plotting.  
```{r echo=FALSE}
library(tidyverse)
library(ggpubr)
data<-read_csv("activity.csv")
```
Let's inspect and get some idea of what the data looks like. 
```{r}
head(data)
summary(data)
```

## What is mean total number of steps taken per day?
We will ignore the NAs in this section. 
We can group the data by date and calculate the mean & median total number of steps taken per day. 
```{r}
totalperday<-data %>%
  group_by(date) %>%
  summarize(`Total Steps`=sum(steps)) %>%
  rename(Date=date)
totalperday
```
Calculate mean and median values. 
```{r}
meanandmedian<-totalperday %>%
  summarize(Mean=mean(`Total Steps`, na.rm=TRUE), Median=median(`Total Steps`, na.rm=TRUE))
meanandmedian
```

Plot a histogram to visualize the total number of steps taken each day.
```{r}
gghistogram(totalperday, x="Total Steps", bins=50)
```

## What is the average daily activity pattern?
We first compute the average number of steps taken across all days for each 5-minute interval.
```{r}
timeseries<-data %>%
  group_by(interval) %>%
  summarise(`Mean Steps`=mean(steps, na.rm=TRUE)) %>%
  rename(Time=interval) %>%
  mutate(Time=format(strptime(sprintf("%04d", Time), format="%H%M"), format="%H:%M"))
timeseries
```
Now we make a time series plot of the 5-minute interval and average number of steps taken, averaged across all days.
```{r}
ggline(timeseries, x="Time", y="Mean Steps")
  
```
The maximum number of steps seem to occur at approximately 8am. 

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
